---
title: "Problem Set 2"
author: "Aubrey Ahern"
date: "11/1/2022"
output: html_document
---

set things up, clear r's memory, import necessary packages
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
#library(ggfortify) may be used in place of library(dplyr) and library(ggplot2)
```

suck in data using here package, look at data using glimpse
```{r}
DF <- read.csv(here("Data", "Halloween_data_PS2.csv"))
glimpse(DF)
```

convert variables costume and type from characters to factors
```{r}
DF$Costume <- as.factor(DF$Costume)
DF$Type <- as.factor(DF$Type)
```

check to ensure the variables were converted correctly by examining data set
```{r}
glimpse(DF)
str(DF)
#str and glimpse are both good functions to use to examine data sets
```

first check to see if there are any errors in the data
```{r}
unique(DF$Costume)
#these are the errors I was able to find...
##error #1 = "Scary", "scary", and "scary " are considered different within the data
##error #2 = "witch", "witch ", and "Witch" are all considered to be different costumes
##error #3 = "Wonder Woman", "Wonder woman", and "wonder woman" are all considered to be different costumes
##error #4 = "Creepy clown", "creepy clown", and "clown" are all considered to be different costumes
```

fix error #1
```{r}
scary <- which(DF$Type == "Scary" | DF$Type == "scary" | DF$Type == "scary ")
DF$Type[scary] <- "Scary"
#the inclusion of a space as well as differences in capitalization cause these variables to be counted separately in R
```

fix error #2
```{r}
witch <- which(DF$Costume == "Witch" | DF$Costume == "witch" | DF$Costume == "witch ")
DF$Costume[witch] <- "witch"
#the inclusion of a space as well as differences in capitalization cause these variables to be counted separately in R
#which vs. witch is confusing lol
```

fix error #3
```{r}
wonder_woman <- which(DF$Costume == "wonder woman" | DF$Costume == "Wonder Woman" | DF$Costume == "Wonder woman")
DF$Costume[wonder_woman] <- "Wonder Woman"
#differences in capitalization cause these variables to be counted separately in R
```

fix error #4
```{r}
clown <- which(DF$Costume == "creepy clown" | DF$Costume == "clown" | DF$Costume == "Creepy clown")
DF$Costume[clown] <- "clown"
#differences in capitalization cause these variables to be counted separately in R
```

calculate mean age of trick-or-treaters and the mean number of pieces of candy taken broken down by costume and type
```{r}
figure1 <- DF %>%
group_by(Costume, Type) %>%
  summarise(
    mean_age = mean(age, na.rm =T),
    mean_candy = mean(candy_pieces, na.rm = T))
#calculated means are stored in my new, figure1 data frame... i can now use this data frame to make a scatterplot...
```

create figure 1
```{r}
ggplot(figure1, aes(x = mean_age, y = mean_candy, color = Type)) +
  geom_point(size = 2) +
  xlab("Mean age of kid (years)") +
  ylab("Mean number pieces taken") +
  theme_bw()
#xlab and ylab indicate axis labels
#this figure looks like the one in the problem set, yay!
#i am curious if we can/will ever include figure captions in R as seen on the problem set
```

isolate the data for children in “not scary” costumes who are 10 years old or younger 
```{r}
figure2 <- DF %>%
group_by(Type = "Not scary") %>%
filter(age <= 10)
#i stored the "not scary" costumes in my new, figure2 data frame... i can now use this data frame to make a histogram as seen in the problemset
```

create figure 2
```{r}
ggplot(figure2, aes(x = age)) +
  geom_histogram(bins = 4) + 
  xlab("Age of child") +
  ylab("Number")
#xlab and ylab indicate axis labels
#set four bars with "bins = 4"
```

(try to) create figure 3
```{r}
ggplot(DF, aes(x = age, y = Costume)) +
  geom_boxplot() +
  geom_point(size = 4, color = 'grey', alpha = 0.5) +
  xlab("Child age (years)") +
  ylab("Costume choice") +
  facet_wrap(~Type)
#reused the data frames i already created for the other questions
#xlab and ylab indicate axis labels
#this looks like the one displayed in the problem set, yay!
```



